---
title: "ESM 244 Lab 3"
author: "Keene Morrow"
date: "1/23/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      message = FALSE,
                      warning = FALSE) # reminder, this is where you can set global options for code chunks

library(tidyverse)
library(janitor)
library(here)
library(sf)
library(tmap)

```

## Binary logistics regression: Penguins!

Read in data:
```{r}
gentoo <- read_csv("gentoo_lter.csv")
chinstrap <- read_csv("chinstrap_lter.csv")

# Use full_join() to merge them together (keeping everything)
penguins <- full_join(chinstrap, gentoo) %>% 
  clean_names() %>% 
  mutate(sex = str_to_lower(sex)) %>% 
  filter(sex %in% c("male","female"))
```

Let's take a look:

```{r}
ggplot(data = penguins, aes(x = sex, y = body_mass_g)) +
  geom_jitter(
    aes(color = sex,
        pch = sex) # changes point style
  ) +
  facet_wrap(~species) +
  theme_minimal()
```

Based on this, it seems pretty likely that we'd be able to identify whether a penguin is a Gentoo or Chinstrap if we know that it's female and 5000g or male and 3000g.
Harder if a 4900g male penguin or 4000g female penguin.



Code
- Chinstrap penguins as 1
- Gentoo penguins as 0
using `case_when`

```{r}
penguin_bin <- penguins %>%
  mutate(
    sp_binary = case_when(
      species == "Chinstrap penguin (Pygoscelis antarctica)" ~ 1,
      species == "Gentoo penguin (Pygoscelis papua)" ~ 0
    )
  )
```

Binary logistic regression
(Attempting to predict species based on sex and body mass)

```{r}
penguin_blr <- glm(sp_binary ~ sex + body_mass_g,
                   family = "binomial",
                   data = penguin_bin)

summary(penguin_blr)
```
Check out the estimates in the summary above.  Remember that the estimates of coefficients are the log odds. The coefficient is the impact on the probability that a penguin is a Chinstrap.

Let's make some predictions for penguins!
(this is pretty close to classification)

1. What is the probability that a penguin is a chinstrap penguin if it weighs 4500g and is male?

```{r}
# models like new data to be presented to them as a shiny new data frame
df_m4500 <- data.frame(sex = "male", body_mass_g = 4500) # variable names and contents must match the original data frame

# find log odds of this penguin being chinstrap:
m4500_logodds <- predict(penguin_blr, newdata = df_m4500, type = "link")

# But was does 4.05 mean?
# Let's make it give us a probability

m4500_logodds <- predict(penguin_blr, newdata = df_m4500, type = "response")

# So, 98.3% chance that it's a Chinstrap penguin
```

2. What is the probability that a penguin is a chinstrap penguin if it weighs 4000g and female?
(this is right on the edge of the two groups)
```{r}
df_f4000 <- data.frame(sex = "female", body_mass_g = 4000)

f4000_logodds <-  predict(penguin_blr, newdata = df_f4000, type = "response")
# 65.4% chance that it's a Chinstrap penguin :)
```

Now some predictions with an entire data frame

```{r}
penguins_mock <- data.frame(
  body_mass_g = rep(seq(3000, 6000, length = 200), 2),
  sex = c(rep("male", 200), rep("female", 200))
)
```

Now let's make lots of predictions

```{r}
full_predict <- predict(penguin_blr, newdata = penguins_mock, type = "response", se.fit = TRUE) # the last bit adds standard error

# Coerce predictions into a data frame, with the mock data + SEs:
final_df <- data.frame(penguins_mock, 
                       full_predict$fit, 
                       full_predict$se.fit)

colnames(final_df) <- c("penguin_mass", "sex", "probability", "se")

# Check out final_df
# Then plot!

ggplot(data = final_df, aes(x = penguin_mass, y = probability)) +
  geom_line(aes(color = sex)) +
  scale_x_continuous(limits = c(3500, 5500)) +
  geom_ribbon(aes(ymin = probability - se,
                  ymax = probability + se,
                  fill = sex),
              alpha = 0.3) +
  labs(x = "Penguin body mass (g)",
       y = "Probability of being Chinstrap") +
  theme_minimal()


```
Machine learning does some of this with the addition of using some data to train the program and reserving the rest of the data to validate the model.

## Now for some spatial data!

```{r}
cougars <- read_sf(dsn = here("cougar_connections"), layer = "ds1014") %>%
  clean_names()
```


```{r}
# sticky geometries demo
large_corriders <- cougars %>%
  filter(area_ac > 20000) %>%
  select(area_ac)

# normally you'd wind up with just area_ac,  but here the geometry sticks around rather than disappear

plot(large_corriders)
```

How about something more interesting
```{r}
cougars_sub <-  cougars %>%
  select(elev_mean)
```

Maps in ggplot:
```{r}
ggplot(data = cougars_sub) +
  geom_sf(aes(fill = elev_mean),
          color = NA)
```

Now add some context with tmap (hey, how about some interactivity too?)
```{r}
tmap_mode("view")

# weird syntax ahead!

tm_shape(cougars_sub) +
  tm_fill("elev_mean") # for polygons

```

