---
title: "ESM 244 Lab 3"
author: "Keene Morrow"
date: "1/23/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      message = FALSE,
                      warning = FALSE) # reminder, this is where you can set global options for code chunks

library(tidyverse)
library(janitor)
library(here)
library(sf)
library(tmap)

```

## Binary logistics regression: Penguins!

Read in data:
```{r}
gentoo <- read_csv("gentoo_lter.csv")
chinstrap <- read_csv("chinstrap_lter.csv")

# Use full_join() to merge them together (keeping everything)
penguins <- full_join(chinstrap, gentoo) %>% 
  clean_names() %>% 
  mutate(sex = str_to_lower(sex)) %>% 
  filter(sex %in% c("male","female"))
```

Let's take a look:

```{r}
ggplot(data = penguins, aes(x = sex, y = body_mass_g)) +
  geom_jitter(
    aes(color = sex,
        pch = sex) # changes point style
  ) +
  facet_wrap(~species) +
  theme_minimal()
```

Based on this, it seems pretty likely that we'd be able to identify whether a penguin is a Gentoo or Chinstrap if we know that it's female and 5000g or male and 3000g.
Harder if a 4900g male penguin or 4000g female penguin.



Code
- Chinstrap penguins as 1
- Gentoo penguins as 0
using `case_when`

```{r}
penguin_bin <- penguins %>%
  mutate(
    sp_binary = case_when(
      species == "Chinstrap penguin (Pygoscelis antarctica)" ~ 1,
      species == "Gentoo penguin (Pygoscelis papua)" ~ 0
    )
  )
```

Binary logistic regression
(Attempting to predict species based on sex and body mass)

```{r}
penguin_blr <- glm(sp_binary ~ sex + body_mass_g,
                   family = "binomial",
                   data = penguin_bin)

summary(penguin_blr)
```
Check out the estimates in the summary above.  Remember that the estimates of coefficients are the log odds. The coefficient is the impact on the probability that a penguin is a Chinstrap.

Let's make some predictions for penguins!
(this is pretty close to classification)

1. What is the probability that a penguin is a chinstrap penguin if it weighs 4500g and is male?

```{r}
# models like new data to be presented to them as a shiny new data frame
df_m4500 <- data.frame(sex = "male", body_mass_g = 4500) # variable names and contents must match the original data frame

# find log odds of this penguin being chinstrap:
m4500_logodds <- predict(penguin_blr, newdata = df_m4500, type = "link")

# But was does 4.05 mean?
# Let's make it give us a probability

m4500_logodds <- predict(penguin_blr, newdata = df_m4500, type = "response")

# So, 98.3% chance that it's a Chinstrap penguin
```

2. What is the probability that a penguin is a chinstrap penguin if it weighs 4000g and female?
(this is right on the edge of the two groups)
```{r}
df_f4000 <- data.frame(sex = "female", body_mass_g = 4000)

f4000_logodds <-  predict(penguin_blr, newdata = df_f4000, type = "response")
# 65.4% chance that it's a Chinstrap penguin :)
```



